Estimating Uncertainty for Tipping Points with Deep Learning

### 2022-04-26 greta/sn_mcmc_with.Rmd
more saddle-node runs:
There are now 2 SN cases: "saddlenode-with" which trains on data with transitions (and forecasts thus start at state 0), and "saddlenode", which attempts the 'true forecast'.

```{r}
set.seed(4242)
train_reps <- 100
train_t_max <- 250
test_t_max <- 250
test_reps <- 100
simulate <- simulate_sn
```

```{r}
p <- list(r = 1,
          K = 1,
          s = 0.1,
          ho = .15,
          alpha = 0.0015/4,
          sigma = 0.02,
          t_init = 0,
          N_init = 0.75
        )
train <- purrr::map_dfr(1:train_reps, 
                      \(i) simulate(t_max=train_t_max, p = p),
                      .id = "i")
test <- purrr::map_dfr(1:test_reps, 
                      \(i) simulate(t_max=test_t_max, p = p),
                      .id = "i")

@@

```{r}
m <- greta_model_sn(train)
bench::bench_time({                 
  draws <- mmcmc(m, n_samples = 700000, warmup = 600000,
                 chains = 6, verbose = FALSE)
})
```

```{r}
## draw test_reps number of samples
combined <- compare_forecast(draws, train, test, simulate, vars = "N",
                              test_reps, test_t_max)
write_csv(combined, "data/stochastic.csv.gz")
```

```{r}
scores <-
  rep_scores(combined, "N") |> 
  mutate(scenario="stochastic", 
         model="MCMC", 
         reps = train_reps)

write_csv(scores, "data/scores_stochastic.csv.gz")

@@

```{r}
combined |> 
  group_by(t,type,variable) |> 
  summarise(mean = mean(value), sd = sd(value), .groups = "drop") |> 
 ggplot(aes(t, col=type)) + 
  geom_ribbon(aes(ymin = pmax(mean-2*sd,0), ymax = mean+2*sd, fill=type), alpha=0.5) +
  geom_line(aes(y=mean)) +
  geom_vline(aes(xintercept = train_t_max)) + facet_wrap(~variable, ncol=1)
```

>>>>

```{r}
set.seed(4242)
train_reps <- 10
train_t_max <- 250
test_t_max <- 250
test_reps <- 100
simulate <- simulate_sn
```

```{r}
p <- list(r = 1,
          K = 1,
          s = 0.1,
          ho = .15,
          alpha = 0.0015/2,
          sigma = 0.02,
          t_init = 0,
          N_init = 0.75
        )
train <- purrr::map_dfr(1:train_reps, 
                      \(i) simulate(t_max=train_t_max, p = p),
                      .id = "i")
test <- purrr::map_dfr(1:test_reps, 
                      \(i) simulate(t_max=test_t_max, p = p),
                      .id = "i")

@@

```{r}
m <- greta_model_sn(train)
bench::bench_time({                 
  draws <- mmcmc(m, n_samples = 10000, warmup = 9000,
                 chains = 6, verbose = FALSE)
})
```

```{r}
## compare to simulations starting at the initial conditions 
# instead of continuing forward from training data
inits <- data.frame(N_init=p$N_init, t_init = 0)
combined <- compare_forecast(draws, train, test, simulate, vars = "N",
                              test_reps, test_t_max, inits = inits
                             )

write_csv(combined, "data/saddlenode_with.csv.gz")
```

```{r}
scores <-
  rep_scores(combined, "N") |> 
  mutate(scenario="saddle-node", 
         model="MCMC", 
         reps = train_reps)

write_csv(scores, "data/scores_saddlenode_with.csv.gz")

@@

```{r}
combined |> 
  group_by(t,type,variable) |> 
  summarise(mean = mean(value), sd = sd(value), .groups = "drop") |> 
 ggplot(aes(t, col=type)) + 
  geom_ribbon(aes(ymin = pmax(mean-2*sd,0), ymax = mean+2*sd, fill=type), alpha=0.5) +
  geom_line(aes(y=mean)) +
  facet_wrap(~variable, ncol=1)
```

### 2022-04-26 manuscript/manuscript.Rmd
more saddle-node runs:
There are now 2 SN cases: "saddlenode-with" which trains on data with transitions (and forecasts thus start at state 0), and "saddlenode", which attempts the 'true forecast'.

---
title: Estimating Uncertainty for Tipping Points with Deep Learning
authors:
  - name: Marcus Lapeyrolerie
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: mlapeyro@berkeley.edu
  - name: Carl Boettiger
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: cboettig@berkeley.edu (corresponding author)
abstract: |
  Lorem Ipsum whatever

@@

>>>>

---
title: Estimating Uncertainty for Tipping Points with Deep Learning
authors:
  - name: Marcus Lapeyrolerie
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: mlapeyro@berkeley.edu
  - name: Carl Boettiger
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: cboettig@berkeley.edu (corresponding author)
abstract: |
  1. Meaningful representations of uncertainty are an essential component of effective ecological forecasting.
  2. Machine learning
  3. We examine scenarios in which stochasticity and non-linearity can lead to extreme uncertainty under common ecological models.  
  4. We show that ..

@@

Uncertainty [@Schindler2015]
